---
title: "Example EFA & CFA Analysis"
output: html_document
author: Fateme Hosseini, Reza Hossini
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library("devtools")
install_github("Reza1317/easyfa")
library("easyfa")
library("data.table")
library("MASS")
```

# Read inventory data
```{r}
data_path = "~/codes/redkite/projects/2020/motivation/data/publication_data/"
fn = paste0(data_path, "/inventory_q25.csv")
df = read.csv(fn)
q_cols = setdiff(colnames(df), c("id"))
df = df
```

# First we set the analysis constants
```{r}
# rotation method for EFA factors
rotate = "varimax"
# filtering threshold for filtering questions / items in EFA
loading_thresh = 0.4
# skewness upper bound for each item
sk_upper = 3
# kurtosis upper bound for each item
kur_upper = 8
# IQR lower bound for each item
iqr_lower = 1
```

# Filter data
## remove ids with too many missings, also remove outliers
```{r}
set.seed(2124)
res = filter_multi_df_outliers_missing(
    df=df,
    value_cols=q_cols)

df = res[["df"]]
res[["missing_df"]]
res[["outlier_df"]][["id"]]
```

# Multivariate normality test (Mardia Coeff)
```{r}
x = df[ , q_cols]
sk = psych::skew(x, na.rm=TRUE, type=3)
sk
kur = psych::kurtosi(x, na.rm=TRUE, type=3)
kur
mardia = psych::mardia(x, na.rm=TRUE, plot=TRUE)
mardia
#plot(1:length(q_cols), abs(sk), col="blue", xaxt="n", ylim=c(0, 8))
#points(1:length(q_cols), abs(kur), col="red")
#axis(1, at=1:length(q_cols), labels=q_cols)
#legend()

sk_kur_df = data.frame(
    "item"=q_cols,
    skewness=round(sk, 2),
    kurtosis=round(kur, 2))

# some questions are rejected by this criteria
# accepted are those which are not rejected
sk_kur_df

summary_df = t(apply(
    X=x,
    FUN=function(x)c(quantile(x, c(0.25, 0.5, 0.75)), "IQR"=IQR(x)),
    MARGIN=2))

# quartiles for the questions
summary_df = cbind(sk_kur_df, summary_df)
summary_df
# this paper says that univariate skewness of the questions should be less than 2 in magnitude
# and the kurtosis should be less than 7 in magnitude
# http://edpsychassociates.com/Papers/EFAguide(2018).pdf
summary_dt = data.table(summary_df)

# we accept the questions which has
# skew less than sk_upper (see above)
# kurtosis less than kur_upper (see above)
# IQR larger than iqr_lower (see above)
summary_dt[ ,
    acceptable := (
        (abs(kur) <= kur_upper) &
        (abs(sk) <= sk_upper) &
        (IQR >= iqr_lower))]

# summary_dt = merge(summary_dt, item_dt, on="item", sort=FALSE)
summary_dt

q_cols_normal = summary_dt[acceptable == TRUE][["item"]]
length(q_cols_normal)
q_cols_normal

# we recalculate the mardia test / qq-plot for the normal qs
df_normal = df[ , as.character(q_cols_normal)]
mardia = psych::mardia(df_normal, na.rm=TRUE, plot=TRUE)
mardia

# small simulation with real normal distribution
# with same number of items and responses
# to see what Mardia coefficient looks like
p = length(q_cols)
n = dim(df)[1]
chol = matrix(runif(p^2)*2-1, ncol=p)
Sigma = t(chol) %*% chol
u = MASS::mvrnorm(n = n, rep(0, p), Sigma, empirical=TRUE)
# mardia = psych::mardia(u, na.rm=TRUE, plot=TRUE)
# reset to normal questions only
# q_cols = q_cols_normal
```

# Divide the data into two parts to perform EFA and CFA
```{r}
half_length = nrow(df) / 2
df_efa = df[1:half_length, ]
df_cfa = df[(half_length + 1):nrow(df), ]
```

# Check EFA assumptions
```{r}
df = df_efa
res = check_efa_assumptions(
    df=df,
    value_cols=q_cols,
    figs_path=NULL)
# bartlet correlation test
res[["barlet_test"]][["p.value"]]
# kaiser sampling adequecy
res[["kaiser_sampling_adequacy"]]
res[["plt_func"]]()
```

# Get number of factors using parallel analysis
```{r}
res = get_factor_num(
    df=df,
    value_cols=q_cols,
    figs_path=NULL)
# R returns: Number of factors = with eigen values > eigen values of random data
factor_num = res[["factor_num"]]
factor_num
```

# Fit the EFA with those number of factors and remove items
```{r}
res = fit_fa_model(
    df=df,
    factor_num=factor_num,
    rotate=rotate,
    value_cols=q_cols,
    loading_thresh=loading_thresh)

df_filtered = res[["df_filtered"]]
# Tucker Lewis Index
res[["tuker_lewis_index"]]
# RMSEA index
res[["rmsea"]]
res[["rejected_items"]]
res[["cfi"]]
accepted_items = res[["accepted_items"]]
```

# Fit the EFA using filtered data
## get clusters plot
```{r}
res = fit_fa_model(
    df=df_filtered,
    factor_num=factor_num,
    value_cols=accepted_items,
    rotate=rotate,
    fm="ml",
    loading_thresh=loading_thresh)

# Tucker Lewis Index
res[["tuker_lewis_index"]]
# RMSEA index
res[["rmsea"]]
# Checking if any item is rejected this time
res[["rejected_items"]]
res[["cfi"]]
factor_item_list = res[["factor_item_list"]]
names(factor_item_list) = paste0("F", 1:length(factor_item_list))

x = factor_item_list

func = function(i) {
    data.frame(
        item=x[[i]],
        factor=rep(names(x)[[i]], length(x[[i]])))
}

cluster_df = do.call(rbind, lapply(X=1:length(x), FUN=func))

fa_model = res[["fa_model"]]
plot(fa_model, cut=loading_thresh)
psych::fa.diagram(fa_model, cut=loading_thresh, cex=1.5)
# total model output from R
fa_model
```


# Calculate MacDonalds' Omega overall and for accepted items
```{r}
omega_model = psych::omega(m=df[ , accepted_items], nfactors=factor_num)
# The ω_h coefficient
omega_model[["omega_h"]]
# limit of omega
omega_model[["omega.lim"]]
# Cronbach Alpha
omega_model[["alpha"]]
# The omega_t coefficient
omega_model[["omega.tot"]]
# The summary statistics for the omega total, omega hierarchical (general) and omega within each group.
omega_model[["omega.group"]]
```

# CFA using lavaan package
```{r}
cluster_dt = data.table(cluster_df)
agg_dt = cluster_dt[ , .(formula=paste0(item, collapse="+")), by=list(factor)]
agg_dt

agg_dt[["formula"]] = paste0(agg_dt[["F"]], "=~", agg_dt[["formula"]])

lavaan_model_formula = paste0(agg_dt[["formula"]], collapse="\n")
cat(lavaan_model_formula)


lavaan_model_formula = ("
Live=~q7+q8+q9+q10+q11+q13
Pos=~q18+q20+q26+q27+q28+q29
Super=~q22+q23+q24+q25
Treat=~q30+q31+q32+q33
Belong=~q35+q36+q37+q39+q40
")


lavaan_cfa_model = lavaan::cfa(lavaan_model_formula, data=df)

summary(lavaan_cfa_model, fit.measures=TRUE, standardize=TRUE)


#Cairo(
#    file=paste0(figs_path, "cfa_sem.png"),
#    type="png",
#    units="in",
#    width=30,
#    height=15,
#    pointsize=5,
#    dpi=72)

semPlot::semPaths(
    lavaan_cfa_model,
    whatLabels="std",
    weighted=TRUE,
    intercepts=TRUE,
    style="lisrel",
    nCharNodes=0,
    nCharEdges=0,
    shapeMan="diamond",
    curveAdjacent=FALSE,
    title=FALSE,
    layout="tree2",
    curvePivot=TRUE)

dev.off()


```
